Discussion

### 2020-03-12 doc/Discussion.rmd
draft work

- DSBs

A. DSB Differences and moderate support for SC length differences
1. More DSBs in the high rec strains
2. The conserved ratios of DSB : CO implies that the increase in DSB numbers are accompanied by proportional increase in crossovers. This is evidence that shifts support away from the CO:NCO decision to an earlier prophase stage such as when the chromatin compaction is set up. 
3. DSB number being driven by the number of chromatin loops which is positively correlated with axis length.
4. These results supported by the moderate support of longer SC length (pachytene stage in the high rec males).
(Review Baier et al 2014)

B.  INTERFERENCE
1. Review results
-The rapid evolution of the genome wide rate in our high rec strains is due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.
-Caveats:  chromosome size effects, 
However, the IFD distributions show less variance in the high rec strains, which would have MORE Chromosomes (Chm1 to Chm7) compared to the low rec strains which have fewer 2Cos (presumably limited to large chromosomes).

-Also:
	-These results are different from the sex differences because the difference is seen in both the raw and normalized IFD.
	-The difference in interference patterns is driven by the difference in the rates of small IFD observations.

@@

# Explanation

@@

>>>>

Our results had two notable instances of evolution in the recombination rate (or correlated trait); evolution of precursor number (DSBs) and crossover interference strength.

## DSBs

Evolution of recombination rates via the number of precursors -- (has been proposed -- ) (Bairer et al). Likely // evidence suggests this pattern is also a feature of the sex differences.

1. More DSBs in the high rec strains
2. The conserved ratios of DSB : CO implies that the increase in DSB numbers are accompanied by proportional increase in crossovers. This is evidence that shifts support away from the CO:NCO decision to an earlier prophase stage such as when the chromatin compaction is set up. 
Together these results indicate that (it's not the CO : NCO decision but an earlier step (when the total number of chromatin loops (the substrate for SPO11, are laid down with the establishment of the meiotic chromosome structure (DNA loop size)

The tethered loop model of DSB formation

3. DSB number being driven by the number of chromatin loops which is positively correlated with axis length.
4. These results supported by the moderate support of longer SC length (pachytene stage in the high rec males).
(Review Baier et al 2014)

_the propotion of DSB repaired into crossovers hasn't evolved, the total number of DSBs has evolved_  (following the tethered lop model of DSBs formation -- this would suggest that in the high rec strains, assuming no large difference in the Mb size of the chromosomes, 
during chromatin remodeling in prophase, in the high rec strains the chromosomes have longer axis, allowing a higher number but shorter DNA loops (for SPO11 to make the DSBs).

## INTERFERENCE evolution

1. Review results
-The rapid evolution of the genome wide rate in our high rec strains is due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.

These results were unexcepted and differ ffrom the major patterns listed in the literature (otto payseur, argawal Korol. (it also goes against the logical assumptions models of fitting more crossovers along the SC chromosome axis space)

Yet (we are confident that this pattern is not due to errors).

 the IFD distributions show less variance in the high rec strains, which would have MORE Chromosomes (Chm1 to Chm7) compared to the low rec strains which have fewer 2Cos (presumably limited to large chromosomes).

-Also:
	-These results are different from the sex differences because the difference is seen in both the raw and normalized IFD, indicateing the mechanism  / physcial strength (interference strength (signal) mediated through the SC has evolved.

	-The difference in interference patterns is driven by the difference in the rates of small IFD observations, (mean vs median). a pattern that is suggestive of selection agisnt short IFDs

There is a model that predicts this positive ocrrelation. The goldstein results (reviewed in ottoPayseur). when the number of crossovers (the obligate crossover assumption is more strict / the
breifly, the model simulates / sets up a scnario where a modifer acts through the probabilty of flanking regions having 0 or 2 crossovers. ith crossover assuance, recombination rate increases the further apart the loci are spaced (?)

Additionally there are a handful of empircal results which also show the unexpected pssotive correlation between gwRR and interference strength.

1. Cattle (Ma?), when comparisons of genetic map length and interference strength are made across strengths within sex, the strain with the higher genome wide recombinaiton rate also has stronger interference (measured as nu gamma parameter)

2. The IFDs from (Beth's cross) a related house mouse cross show that the PWD (same strain, different mice) had longer IFD compared to Cast (HVR unpublished data). (also gamma?)

3. In Peromyscus, the wild mice had higher gwRR and stronger (gamma nu estimates) compared to lab raised mice of the same speicies [@Peterson]

<note about the differences in the patterns>  assumptions with gamma dstribution and linkage maps, (closer to infinte genetic space), less weight on crossover assurance - more weigth on 
(pooled SC/chromosome length gives strong correlation )

<which is much different than when a single chromosome is examined (by itself), finite 2D length, crossover assurance, weaker correlation with ((the closer to chromosome level for SC legnth metric / summary, the weaker the correlation with crossover number))

@@

# Explanations and Model testing

This section refers to the matrix of models and results from this study where we (conceptually test if each model would support our results)

We were motivated to run a comparison of

@@

be aware of sex-specific and sex average biases

- linkage / R between two foci is sex averaged

- Celegans -- cytlogy focuses on oogenesis  and genetic maps are often from hermaphrodites (selfing -- sperm and egg landscapes are merged in hemaphordites)

-- the value of Chromosome level measures -- incorporate chromosome behavior into models

-  more attention to cell physiology

-- lack of SC length prophase cell charaecterization

-- comparisons within and between sexes (that cattle paper)
